---
title: "Practical Machine Learning Project"
author: "Vu Nguyen"
date: "August 24, 2014"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

Data preprocessing
----------------

We load the training and testing data into variables `Train` and `Test`.
```{r, message=FALSE}
library(caret)
Train = read.csv("pml-training.csv")
Test = read.csv("pml-testing.csv")
```

Upon `str(Train)` inspection, we find that many columns (out of 160) contain mostly `NA`s. We remove these columns in the first pass of cleaning process.

```{r}
nNAs = apply(Train, 2, function(col) {
  sum(is.na(col) | col == "")
})
Train = Train[ , -which(nNAs > 0.9 * length(Train$classe))]
```

We also remove columns which have near zero variance, i.e. they don't contribute to the model.
```{r}
Train = Train[ , -nearZeroVar(Train)]
```

In the final pass, we will remove columns that has nothing to do with the model, such as `user_name`, `raw_timestamp_part_1`, etc.

```{r}
Train = Train[ , !names(Train) %in% c("X", "user_name", "raw_timestamp_part_1", "raw_timestamp_part_2", "cvtd_timestamp", "new_window", "num_window")]
dim(Train)
```

We are now left with tidy data, ready to build some models.

Model Construction
------------
We will construct a Random Forests (rf) model. It's relatively simple to set this up and have it running. We first partition our `Train` data into `training` and `testing` to verify out-of-sample error.

```{r}
trainIndex = createDataPartition(Train$classe, p = 0.75, list = F)
training = Train[trainIndex, ]
testing = Train[-trainIndex, ]
fitCtrl = trainControl(method="cv", number = 5, verboseIter = F)
rfFit = train(classe ~ ., data=training, method="rf", trControl = fitCtrl, verbose=F)
rfFit
```

Confusion matrix on the `testing` set. The result is great.

```{r}
confusionMatrix(predict(rfFit, testing), testing$classe)
```

Finally, we run `rfFit` on `Test`, obtaining the following result.
```{r}
predict(rfFit, Test)
```

Reference
------------
Ugulino, W.; Cardador, D.; Vega, K.; Velloso, E.; Milidiu, R.; Fuks, H. Wearable Computing: Accelerometers' Data Classification of Body Postures and Movements. Proceedings of 21st Brazilian Symposium on Artificial Intelligence. Advances in Artificial Intelligence - SBIA 2012. In: Lecture Notes in Computer Science. , pp. 52-61. Curitiba, PR: Springer Berlin / Heidelberg, 2012. ISBN 978-3-642-34458-9. DOI: 10.1007/978-3-642-34459-6_6.